Proof No Diamond (3 layers)

### 2024-01-08 Paper/Paper.tex
Draft completed but need big revision.

@@

\begin{align*}
	\mathrm{UB}_{a^0_J,a^0_I} = B_z + \sum_{b\in B_O} w_{b,z}\ReLU(weight_{a^0_I,a^0_J}(b)) + \sum_{b\in B_{pure}} w_{b,z} \mathrm{AppB}(weight_{a^0_I,a^0_J}(b)).
\end{align*}

where $\mathrm{AppB}(weight_{a^0_I,a^0_J}(b))$ is the upper bound of 
the (LP) approximation of $weight_{a^0_I,a^0_J}(b)$ if $w_{b,z}>0$, and the lower bound
of its (LP) approximation if $w_{b,z}<0$.

@@

$$\sum_{b\in B^2} w_{b,z} \mathrm{AppB}(weight_{a^*_I,a^0_J}(b)) = 
\sum_{b\in B^2} w_{b,z} \mathrm{AppB}(weight_{a^*_I}(b))
=\sum_{b\in B^2} w_{b,z} \mathrm{ReLU}(weight_{a^*_I}(b))
$$

as $\mathrm{AppB}(weight_{a^*_I}(b))=\mathrm{ReLU}(weight_{a^*_I}(b))$ since 
$weight_{a^*_I}(b)$ is an extremal value (minimum or maximal) and 
$\mathrm{AppB}$ equals $\mathrm{ReLU}$ for extremal values (make a lemma about that in the begining).

@@

$$\mathrm{UB} = max_{a_J,a_I} \mathrm{UB}_{a_J,a_I} \leq \max_{a_J,a_I} weight_{a_J,a_I}(z)$$

@@

%
%The DeepPoly function for $b$ can be expressed by the following steps.
%First, we partition the set of indices $i$ as $I,I_0,I^+,I^-$ with:
%\begin{itemize}
%	\item $i \in I$ if the sign of $b_i^-=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$ is always positive irrespective of $Y=(a_j)_j$, in which case 
%	$\bar{f}(b_i^+)(Y)=\underline{f}(b_i^+)=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$, and else:
%	\item $i \in I_0$ if the sign of $b_i^-=\sum_{j} w_{a_j b_i} a_j + B_{b_i}$ is always negative irrespective of $Y=(a_j)_j$, in which case $\bar{f}(b_i^+)=\underline{f}(b_i^+)=0$,	and else:
%	\item $i \in I^+$ if $w_{b_i c}>0$, and:
%	\item $i \in I^-$ if $w_{b_i c}<0$.
%	\end{itemize}
%	
%Remark that for all $i \notin I$, we have $\underline{f_i}(b_i^+)(Y)=0$.
%	
%
%We have:
%\begin{align*}
%	\bar{f}(c)(Y) = &
%	bias_c + \sum_{b \in I} w_{bc}\sum_{a} w_{a b} weight(a) + bias_{b} \\
%	& + \sum_{b \in I^+} w_{b c} \cdot \bar{f}(ReLU(\sum_{a} w_{a b} weight(a) 
%	+ Bias_{b}))(Y) 
%	\end{align*}

>>>>

\begin{lemma}\label{lem:AppB=ReLU}
	Let $\mathrm{AppB}$ be the upper bound of lower bound approximation functions of LP for $\ReLU$ function with known upper and lower bound (numbers) $u,l$. Then when $x$ reaches the extremal values (maximum or minimum values, $u$ or $l$), then $\mathrm{AppB}(x)=\ReLU(x)$.

\end{lemma}

\begin{proof}
	This is simply by definition of LP approximation functions.
\end{proof}

@@

\begin{align*}
	\mathrm{UB}_{a^0_J,a^0_I} = B_z + \sum_{b\in B_O} w_{b,z}\ReLU(weight_{a^0_I,a^0_J}(b)) + \sum_{b\in B_{pure}} w_{b,z} \mathrm{AppB}(weight_{a^0_I,a^0_J}(b)).
\end{align*} where $\mathrm{AppB}(weight_{a^0_I,a^0_J}(b))$ is the upper bound of 
the (LP) approximation of $weight_{a^0_I,a^0_J}(b)$ if $w_{b,z}>0$, and the lower bound
of its (LP) approximation if $w_{b,z}<0$.

@@

$$\sum_{b\in B^2} w_{b,z} \mathrm{AppB}(weight_{a^*_I,a^0_J}(b)) = 
\sum_{b\in B^2} w_{b,z} \mathrm{AppB}(weight_{a^*_I}(b))
=\sum_{b\in B^2} w_{b,z} \mathrm{ReLU}(weight_{a^*_I}(b))
$$ as $\mathrm{AppB}(weight_{a^*_I}(b))=\mathrm{ReLU}(weight_{a^*_I}(b))$ since 
$weight_{a^*_I}(b)$ is an extremal value (minimum or maximal) and 
$\mathrm{AppB}$ equals $\mathrm{ReLU}$ for extremal values.

@@

$$\mathrm{UB} = max_{a_J,a_I} \mathrm{UB}_{a_J,a_I} \leq \max_{a_J,a_I} weight_{a_J,a_I}(z)$$ And this us what we want to show.

\section{Proof of Theorem 2, full version}

In this section, we will try to prove the full version of theorem 2, that is

\begin{theorem}
	\label{no_diamond_3}

	Suppose we have a DNN with $\ReLU$ as the unique activation functions. We consider compensating pairs in between any two nodes (i.d., the source node or target node can in hidden layers.)

	Using MILP method, if every nodes $b$ in any compensating pair
	$(\pi,\pi')$ is encoded as a binary/integer variable, then the upper and lower 
	bounds computed by MILP are the exact max and min value of $z$.
\end{theorem}

Before the proof of above theorem, we will try to weaker one, that is, consider not only compensating pairs, but also pairs that allows common nodes (the two paths in a compensating pairs cannot have a common nodes). We call such pair a weak compensating pair.

We uses very similar definitions as in previous section.

\begin{definition}
	Let $K$ be the set of all input nodes $a_k$. 
	We define a decomposition $K=I\sqcup J$ as follows:  
	\begin{itemize}
		\item $k \in I$  if
		\begin{enumerate}
			\item every path from $a_k$ to $z$ has weight $\geq 0$, or
			\item every path from $a_k$ to $z$ has weight $\leq 0$.
		\end{enumerate}
		That is, $a_k$ is not a source of a weak compensating pair.
		\item $k \in J$ if $k \notin I$, that is there exists two paths $\pi,\pi'$ from $a_k$, 
		one with positive and one with negative weight.
	\end{itemize}
\end{definition}

\begin{lemma} \label{lem:open_node_2}
	A node $b$ in the hidden layer will not be on a weak compensating pair iff one of the following two happens:
	\begin{enumerate}
		\item $w_{b,z}=0$, or
		\item For every input node $j\in J$, we have $w_{a_j,b}=0$.
	\end{enumerate}
	We denote $B_{pure}$ the set of such nodes $b$ such that at least one of the above holds.
\end{lemma}

\begin{proof}
	First, we show that if either one of 1,2 happens, then $b$ will not be opened. If 1), it is obvious. For 2, we reason by contradiction: assume there is a pair of weak compensating paths 	$(\pi,\pi')$ starting with $a$, with $b$ in $\pi$ and weight$(\pi) > 0$. It means that $a \in J$. A contradiction as 2) $w_{a_j,b}=0$ implies weight$(\pi)=0$.

	Second, we show that if neither 1 nor 2 hold, then $b$ will be on a weak compensating path.
	Because 2) does not hold, there is a $j \in J$ with $w_{a_j,b} \neq 0$, say $>0$.
	Because 1) does not hold, $w_{b,z} \neq 0$, say $>0$.
	Now, by definition of $J$, there is a pair of weak compensating paths $\pi,\pi'$ 
	from $a_j$ to $z$, say with $\pi'$ with weight $<0$.
	Then $((a_j,b,z), \pi')$ is also a weak compensating pair.
\end{proof}

\begin{definition}\label{sign_of_nodes_in_I_2}
	We define a partial sign function $S$ over nodes $n$ such that : 	
	\begin{enumerate} 
		\item all paths from $n$ to $z$ have 0 weight, and then $S(n)=0$; 
		\item all paths from $n$ to $z$ have non-negative weight, and at least one path has a positive weight, and then $S(n)=1$; 
		\item all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight, and then $S(n)=-1$.
		\item Otherwise $S(n)$ is undefined.
	\end{enumerate}
\end{definition}

\begin{lemma} \label{lem:reach_max_2}
	$max_a (weight_{a}(z)) = max_{\{a \mid a_I=a^*_I\}} (weight_{a}(z))$

	Now, for any fixed input $a^0_J$, we can regard $weight_{a_I,a^0_J}(z)$ as a DNN $D'$ with input nodes $a_i\in I$, with all $a^0_J$ %and their propagation in hidden layers 
	as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply Theorem \ref{th1} to get that $z(\cdot,a^0_J)$ reaches its maximal value for $a_I=a_I^*$, and we are done.

	For the second part, for any $b$ that is at least 2 layers after $a$, we can regard $b$ as the output node and apply the same argument above.

	If $b$ is immediately after $a$, then the proof is the same as last part of Lemma \ref{lem:reach_max} last section.

\end{proof}

Denote by $\mathrm{UB}$ the maximal bound on $z$ and $\mathrm{LB}$ the minimal bound on $z$ considering the MILP constraints above, with ReLUs being either linear or binary depending on whether $b \in B_{pure}$ or not.

For an input vector $a^0_I,a^0_J$, let $UB_{a^0_I,a^0_J}$ be the upper bound in the fixed formulation with input $a^0_I,a^0_J$, and similarly for $LB$ lower bound. We have the following lemma:

\begin{lemma} In an MILP formulation, if $c$ is a node:

	1. $UB(b)=\max_{a^0_I,a^0_J}UB_{a^0_I,a^0_J}(c)$.

		1. $LB(b)=\min_{a^0_I,a^0_J}LB_{a^0_I,a^0_J}(c)$. 
\end{lemma}

\begin{proof}
	This is by the definition of MILP formulation and $UB_{a^0_I,a^0_J}, LB_{a^0_I,a^0_J}$.
\end{proof}

\begin{lemma}\label{lem:pure_node}
	For a node $c\in B_{pure}$ such that $w_{b,z}\neq 0$, $UB_{a^0_I,a^0_J}$ (and $LB$ the same) only depends on $a^0_I$.
\end{lemma}

\begin{proof}
	We prove this lemma by induction. For the first hidden layer, this is trivial.

	Suppose for layer up to $l_i$, this lemma is true, we show it for layer $l_{i+1}$. For a node $c$ in layer $l^{i+1}$, we have that: \begin{align*}
		UB_{a^0_I,a^0_J}(c) = B_c + \sum_{b\in B_O\cap l_i} w_{b,c}\ReLU(B_{a^0_I,a^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}} w_{b,c} \mathrm{AppB}(B_{a^0_I,a^0_J}(b)).
	\end{align*} But since $c\in B_{pure}$, for any $b\in B_O$, we must have $w_{b,c}=0$. Therefore, we will have:

	\begin{align*}
		UB_{a^0_I,a^0_J}(c) = B_c + \sum_{b\in B_{pure}\cap l_{i}} w_{b,c} \mathrm{AppB}(B_{a^0_I,a^0_J}(b)).
	\end{align*} By induction, for any $b\in B_{pure}\cap l^i$, either $w_{b,z}=0$ or it only depends on $a^0_I$. Since $w_{c,z}\neq 0$, if $w_{b,z}=0$, we must have $w_{b,c}=0$. Therefore, we will have:

	\begin{align*}
		UB_{a^0_I,a^0_J}(c) = B_c + \sum_{b\in B_{pure}\cap l_{i}\wedge w_{b,c}\neq0} w_{b,c} \mathrm{AppB}(B_{a^0_I}(b)).
	\end{align*} So, $UB_{a^0_I,a^0_J}(c)$ only depends on $a_I^0$.
\end{proof}

We want to show that $\mathrm{UB}(z) = \max z$ and $\mathrm{LB}(z) = \min z$ for the unique output nodes and we will prove this for all $c$ such $w_{c,z}\neq 0$. As the MILP abstraction is a sound overapproximation, 
it suffices to show that $\mathrm{UB}\leq \max c$ and $\mathrm{LB} \geq \min c$. We will prove this by induction on layers from the first hidden layer to the output layer. More specifically, we will prove the following lemma:

\begin{lemma}
	For the $i$-th layer $l^i$ ($1$-th layer is the first hidden layer), for a node $c\in l^i$, if $w_{c,z}\neq 0$, then:

	 For any fixed $a^0_J$, $\max_{a_I} UB_{a_I,a^0_J}(c)=UB_{a^*_I,a^0_J}(c)= weigth(a^*_I,a^0_J)(c) = \max_{a_I} weigth(a_I,a^0_J)(c)$, where $a^*_I$ is described in previous section.

The similar is true for lower bound $LB$.
\end{lemma}

\begin{proof}

	For $i\leq 2$, we have proved in previous section.

	Suppose we have proved this lemma up to layer $i\geq 2$, and we are going to prove it for $i+1$-th layer. Let $c$ be a node on $i+1$-th layer such that $w_{c,z}\neq 0$.

	By definition, we have that for any input vector $a^0_I,a^0_J$:

	\begin{align*}
		\mathrm{UB}_{a^0_J,a^0_I}(c) = B_c + \sum_{b\in B_O\cap l_i} w_{b,c}\ReLU(B_{a^0_I,a^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}} w_{b,c} \mathrm{AppB}(B_{a^0_I,a^0_J}(b)).
	\end{align*} where $\mathrm{AppB}$ is the same as in previous section  whose definition is independent with $a^0_I,a^0_J$. $B$ is one of $UB,LB$  that, if $w_{b,c}>0$, then $B$ is $UB$ and otherwise $B$ is $LB$.

	By Lemma \ref{lem:pure_node}, we can simplify above equation to \begin{align*}
			\mathrm{UB}_{a^0_J,a^0_I}(c) = B_c + \sum_{b\in B_O\cap l_i} w_{b,c}\ReLU(B_{a^0_I,a^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}\wedge w_{b,c}\neq 0} w_{b,c} \mathrm{AppB}(B_{a^0_I}(b)).
	\end{align*}

	By induction hypothesis, since both $\ReLU$ and $\mathrm{AppB}$ are non-decreasing functions, we have that \begin{align*}
		\mathrm{UB}_{a^0_J,a^0_I}(c)\leq 
		 &B_c + \sum_{b\in B_O\cap l_i} w_{b,c}\ReLU(B_{a^*_I,a^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}\wedge w_{b,c}\neq 0} w_{b,c} \mathrm{AppB}(B_{a^*_I}(b))\\
		= & \mathrm{UB}_{a^0_J,a^*_I}(c) \\
	\end{align*}

Now, since when $x$ reaches the max or min of $\mathrm{AppB}$'s domain, $\mathrm{AppB}(x)=\ReLU(x)$, and by induction hypothesis part 2, for any $b\in B_{pure}\cap l_i\wedge w_{b,c}\neq 0$, $\mathrm{AppB}(B_{a^*_I}(b))=\ReLU(B_{a^*_I})$. Hence, by induction hypothesis, we will have \begin{align*}
	\mathrm{UB}_{a^0_J,a^0_I}(c)\leq & \ \mathrm{UB}_{a^0_J,a^*_I}(c) \\
	= &B_c + \sum_{b\in B_O\cap l_i} w_{b,c}\ReLU(B_{a^*_I,a^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}\wedge w_{b,c}\neq 0} w_{b,c} \mathrm{AppB}(B_{a^*_I}(b))\\
	= &B_c + \sum_{b\in B_O\cap l_i} w_{b,c}\ReLU(weight_{a^*_I,a^0_J}(b)) + \sum_{b\in B_{pure}\cap l_{i}} w_{b,c} \mathrm{ReLU}(weight_{a^*_I}(b))\\
	= & weight_{a^*_I,a^0_J}(c).
\end{align*} This finishes the proof.
	\end{proof}

Now, we have that for any fixed $a^0_J$, $\max_{a_I} UB_{a_I,a^0_J}(c)= \max_{a_I} weigth(a_I,a^0_J)(c)$, therefore, $\max UB_{a_I,a_J}(c)= \max weigth(a_I,a_J)(c)$ and this is to say $UB(c)=\max(c)$.

@@
